Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-05-13 Nicolas.tex
NN changes

\usepackage[dvipsnames]{xcolor}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
\newcommand{\nnn}[1]{{\textcolor[rgb]{0.40, 0.0, 0.20}{#1}}}
\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
\newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}

@@

\author{
    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
    \and \mr{M. Rigault} \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
    \orcid{0000-0002-8121-2560}
    \and\\ \yc{Y. Copin} \inst{1}
    \orcid{0000-0002-5317-7518}
    \and R. Graziani \inst{2}
    %HERE ALPHABETIC ORDER
    \and G. Aldering\inst{3}
    \and M. Briday\inst{1}
    \and J. Nordin\inst{4}
    \and A.~G. Kim\inst{3}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    % Gerard ne souhaite pas etre co-auteur.
}

@@

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    redshift drift. In this paper, we study the redshift evolution of SN~Ia
    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a redshift evolution?}} The SN stretch has been shown to strongly
    correlate with the SN environment, notably with stellar age tracers. We
    model the underlying stretch distribution as a function of redshift, using
    the evolution of fraction of young and old SNe~Ia as predicted by
    \cite{rigault2018}, and assuming constant underlying stretch distribution
    for each age population made of Gaussian mixtures.  We test our prediction
    against \yc{published samples} chosen to have negligible \yc{magnitude} selection effects,
    \yc{so} that any observed \yc{change} is indeed \yc{of astrophysical and not
    observational origin}. We clearly demonstrate that the underlying SN~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a much better description of the data than any
    \yc{\sout{non-drifting} constant} model, including the sample-based asymmetric distributions
    usually used to correct Malmquist bias. \yc{The} favored underlying stretch model
    is \yc{the bimodal one introduced in \cite{rigault2018}}: a high-stretch mode shared by both young and old environments,
    and a low-stretch mode exclusive to old environments. The precise impact of
    the redshift evolution of the SN Ia population \yc{intrinsic properties} on cosmology
    remains to be studied. Yet, the astrophysical drift of the SN stretch
    distribution does affect current Malmquist bias corrections and thereby
    distances derived from SN affected by selection effects. We highlight that
    such a bias will increase with surveys covering \yc{increasingly} larger redshift ranges,
    which is particularly important for LSST.}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. \mr{These 114 SNfactory SNe~Ia are thus \yc{in} the volume limited
part of the survey (Aldering \yc{et al.}, in prep.)}, \yc{and are therefore} assumed to be
a random sampling of the underlying SN population. The SNfactory sample is
particularly useful for studying SN property drift, as it enables us to have a
large SN~Ia sample at $z<0.1$.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{multline}\label{eq:stretchz}
    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
\end{multline}

@@

Given the probability $y^i$ that a given SN is young, and assuming our Base
\nn{model} (see Section~\ref{sec:basemodel}), the probability to measure a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
\end{align}

@@

\mr{Following \cite{rigault2018}, we \yc{define?} $p_y^i$ as one minus the cumulative
distribution function of the $\log(\mathrm{LsSFR})$ at
$-10.82\,\mathrm{dex}$. This is defining the probability that a SN progenitor
is classified as young (see further discussion about this assumption \yc{\emph{which assumption, if it's purely your own definition of young vs. old?}} in Briday
et al. in~prep). Since the drift is negligible within the SNfactory dataset
($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset),
one can readily use $y^i = p_y^i$ to minimize Eq.~\ref{eq:likelihood} with
respect to $\vec{\theta}$. Results on fitting the SNfactory SNe with this model
are shown Table~\ref{tab:modelresults} and illustrated in
Fig.~\ref{fig:stretchlssfr}.} \yc{\emph{Now I'm confused why one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is available, but you seem to say it's an assumption.}}

@@

%\agk{You don't define $p^i_y$. The language is is sloppy. What is relevant is
%the opposite, the probability that a galaxy with low mass gets the measured
%LsSFR. These are in general different. The solution you use is only correct
%for a model in which the rates for low and high-LsSFR galaxies is the same.
%The probability in general should include a prior on $l$.

%\begin{align}
%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
% p(l_o^i | li) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
% p(l_o^i | l^i) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
% p(l_o^i | l^i) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
%\end{align}
%Note that you use rates without LsSFR measurements, so you are doing something
%inconsistent here. There should be a $\delta{z}$ term here. However, since the
%rates are the same at $z=0.05$, maybe this doesn't matter so much.
%}

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances}.
We see in this figure that the measured mean SN~Ia stretch per redshift bins of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-hist_on.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift. Markers show the mean stretch measured in
        redshift bins of equal sample size, indicated in light gray at the
        bottom of each redshift bins. Full and light markers are used when
        considering the fiducial or the conservative samples, respectively. The
        orange horizontal line represents the mean stretch of the fiducial
        sample, illustrating the expectation if the SN stretch distribution is
        not evolving with redshift. Best fits of our Base drifting model are
        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
        the conservative one or the SNfactory dataset only, respectively; all
        are compatible. The light-blue band illustrates the amplitude of the
        error (incl. covariance) of the best fit model when considering the
        fiducial dataset. \mr{The redshift distributions of each sample are
        shown as light-histograms, see lower left legend.} \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
    \label{fig:modelall}
\end{figure*}

@@

Since the various models have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes extra degrees of
freedom to avoid over-fitting the data, and is defined as follow:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k
\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model is the one
with the smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
a Gaussian probability distribution. \mr{However, we highlight that the best of
the compared models might not be the true underlying model.} \yc{\emph{why this sentence? Is there any \textbf{true} intrinsic model?}}

@@

\yc{Remark that, when comparing models adjusted on individual subsamples rather than globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
k\ln(n)$, with $n$ the number of data points) might be better suited than AIC, since it explicitly accounts for
the fact that each subsample is fitted separately: the sample-based model BIC is
rightfully the sum of the BIC for each sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, 
again \sout{excluding} \emph{refuting} the sample-based asymmetric Gaussian model as being as pertinent as the Base model.}

%\agk{I think it does worse not because of the more parameters but that it has 5
%complexity penalties whereas the base model only has 1.  AIC doesn't account
%for the fact that each sample is fit independently.  It seems a fairer metric
%would be the BIC, where you adjust $n$ for each sample.  This would recover
%$\sim 16$ in the IC for this model.}

@@

The amplitude of this magnitude bias for cosmology is beyond the scope of this
paper given the complexity of the BBC analysis. It would require a full study
using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
asymmetric modeling as part of the BBC simulations.
\mr{However, we already highlight that even if a non-drifting sample-based model
could provide comparable result in the volume-limited part of the various
samples, these models would differ when extrapolating at higher redshifts,
precisely where the  underlying \yc{\emph{missing word?}} will matter for correcting Malmquist biases.}

%Nonetheless, it is enlightening to study the expected mean stretch difference
%as a function of redshift when considering that the underlying SN stretch
%distribution follows either the per-sample Asymmetric Gaussian modeling or our
%Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
%we convert this difference in mean stretch $x_1$ to a difference in magnitude
%using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
%correspond to the actual impact on cosmological distances derived from SNe~Ia
%as this simple stretch average comparison does not account for the amplitude of
%the Malmquist bias correction to be applied as a function of redshift and
%surveys.

%\agk{ The AG modeling contains Malmquist bias. To be fair you should plot Base
%with the survey-specific Malmquist bias. There is something strange going on at
%$z<0.2$, can you add SNfactory to this plot? I think you fill find that the
%systematic difference is driven by differences in the $x_1$ distribution of
%SNf.}

%\begin{figure}
%    \centering
%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
%    \caption{Difference of mean stretch standardization (in mag, using
%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
%        Gaussian asymmetric modeling and our Base drifting model, as a function
%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
%        considering the individual SN redshifts (no redshift cut applied). The
%        orange markers represent their mean evolution in 30 regularly-spaced
%        redshift bins. This must be multiplied by the amplitude of the Malmquist
%        bias correction to derive the SN distance modulus bias affecting
%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
%        the mean of all surveys combined. See Greg's comments.}}}
%    \label{fig:magdrift}
%\end{figure}

%We see that the amplitude of the modeling difference is a few tens of millimag.
In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
stress that correctly modeling of potential SN redshift drift should be further
studied and care should be made when using samples affected by selection
effects.

@@

In addition to this ``base'' modeling, we have studied various distributions,
including redshift independent models; we also studied the prediction from the
``Beams with Bias Correction'', a Malmquist bias correction algorithm used by
several recent SN~Ia cosmological analyses, implemented as a per-sample
asymmetric Gaussian distribution. \yc{\emph{of what?}}

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is \yc{significantly} redshift dependent, as
        previously suggested by e.g.~\cite{howell2007}. This result is largely
        independent of details on each age-population modeling.

    \item \yc{\sout{Non-drifting} Redshift-independent} models are indeed excluded as suitable descriptions of
        the data relative to our Base model. This model assumes that: (1) the
        younger population has a unimodal Gaussian stretch distribution, while
        the older population stretch distribution is bimodal, one mode being the
        same as the young one; (2) the evolution of the relative fraction of
        younger and older SNe~Ia follows the prediction made in
        \cite{rigault2018}. This second result strongly supports the existence
        of both young and old SN~Ia populations, in agreement with rate studies. \yc{\emph{ref???}}

    \item Using survey-based asymmetric Gaussian distributions, as done, e.g.,
        in the current implementation of the BBC, is excluded to be good
        descriptions \yc{\emph{rephrase: "using is excluded"?}} of the data relative to our drifting model. Hence, the
        sample-based approach does not accurately account for redshift drift and
        even less so as survey span \yc{increasingly} larger redshift ranges. We \yc{\sout{warn} stipulate} that, even
        if extra degrees of freedom might be acceptable given the large
        number of SNe~Ia in cosmological studies, extrapolating the SN property
        distributions from the volume-limited part of a survey to its
        \yc{Malmquist-biased} magnitude-limited one \yc{\sout{affected by sample selection}} would still be
        inaccurate \yc{\sout{since the distribution would drift} because of the redshift evolution}.

    %\item An inaccurate underlying stretch distribution model is estimated to
    %    bias the mean standardized magnitude of SNe~Ia affected by
    %    selection effects at the order of a few percents. This is
    %    comparable in scale to the observational signature of %exotic forms of
    %    dark energy and is therefore a significant source of systematic that
    %    should be carefully studied.

    \item Given the current dataset, we suggest the use of the following stretch
        population model as a function of redshift:
        \begin{align*}
        \label{eqconclusion:stretchz}
            X_1\left(z \right) =
             &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
            (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
            (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
            \tag{\ref{eq:stretchz}}
        \end{align*}
        with $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
        $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}), and using the
        age-population drift model         \begin{align*}
            \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
            \tag{\ref{eq:delta}}
        \end{align*}
        with $K=0.87$.
\end{enumerate}

In this paper, we considered a simple Gaussian mixture modeling, but additional
data free from significant Malmquist bias would enable us to refine it \yc{as necessary}. We note
that samples at the low- and high-redshift ends of the Hubble diagram would be
particularly helpful for this drifting analysis; fortunately this will soon be
provided by the Zwicky Transient Facility \citep{bellm2019, graham2019} and
Subaru and SeeChange SNe~Ia programs, respectively. \yc{\emph{cannot say who's who in 'A and B and C'.}}

>>>>

\usepackage[dvipsnames]{xcolor}
\newcommand{\mr}[1]{\textcolor[rgb]{0.60,0.10,0.6}{#1}}
\newcommand{\mri}[1]{\textcolor{red}{#1}}
\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
\newcommand{\nnn}[1]{\nn{\textit{#1}}}
\newcommand{\yc}[1]{\textcolor{BrickRed}{#1}}
\newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}

@@

\author{
    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
    \orcid{0000-0002-8121-2560}
    \and\\ Y. Copin \inst{1}
    \orcid{0000-0002-5317-7518}
    \and R. Graziani \inst{2}
    %HERE ALPHABETIC ORDER
    \and G. Aldering\inst{3}
    \and M. Briday\inst{1}
    \and J. Nordin\inst{4}
    \and A.~G. Kim\inst{3}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    % Gerard ne souhaite pas etre co-auteur.
}

@@

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    redshift drift. In this paper, we study the redshift evolution of SN~Ia
    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
    property, to probe this drift. \yc{\emph{isn't a 'drift' \textbf{exactly} a
    redshift evolution?}} The SN stretch has been shown to strongly correlate
    with the SN environment, notably with stellar age tracers. We model the
    underlying stretch distribution as a function of redshift, using the
    evolution of fraction of young and old SNe~Ia as predicted by
    \cite{rigault2018}, and assuming constant underlying stretch distribution
    for each age population made of Gaussian mixtures. We test our prediction
    against published samples chosen to have negligible magnitude selection
    effects, so that any observed change is indeed of astrophysical and not
    observational origin. We clearly demonstrate that the underlying SN~Ia
    stretch distribution is evolving as a function of redshift, and that the
    young/old drifting model is a much better description of the data than any
    \yc{\sout{non-drifting} constant} model, including the sample-based
    asymmetric distributions usually used to correct Malmquist bias. The favored
    underlying stretch model is \yc{the bimodal one introduced in
    \cite{rigault2018}} \nnn{to my knowledge, R+18 doesn't introduce a
    stretch model, but rather discusses the differences between the stretch
    distributions for the young and old pop, never saying anything about stretch
    modes}: a high-stretch mode shared by both young and old environments, and a
    low-stretch mode exclusive to old environments. The precise impact of the
    redshift evolution of the SN Ia population intrinsic properties on
    cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
    distribution does affect current Malmquist bias corrections and thereby
    distances derived from SN affected by selection effects. We highlight that such
    a bias will increase with surveys covering increasingly larger redshift
    ranges, which is particularly important for LSST.}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. These 114 SNfactory SNe~Ia are thus in the volume limited part
of the survey (Aldering et al., in prep.), and are therefore assumed to be a
random sampling of the underlying SN population. The SNfactory sample is
particularly useful for studying SN property drift, as it enables us to have a
large SN~Ia sample at $z<0.1$.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{align}\label{eq:stretchz}
    X_1(z) = \delta(z)&\times \mathcal{N}(\mu_1,\sigma_1{}^2) + \nonumber \\
    (1-\delta(z))&\times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right]
\end{align}

@@

Given the probability $y^i$ that a given SN is young, and assuming our Base
model (see Section~\ref{sec:basemodel}), the probability to measure a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
\end{align}

@@

Following \cite{rigault2018}, we define $p_y^i$ as one minus the cumulative
distribution function of the $\log(\mathrm{LsSFR})$ at $-10.82\,\mathrm{dex}$.
This is defining the probability that a SN progenitor is classified as
young\nn{, thus assuming the LsSFR to be a perfect age tracer} (see further
discussion about this assumption \yc{\emph{which assumption, if it's purely your
own definition of young vs.\ old?}} in Briday et al.\ in~prep). \nn{This
definition should be weighted by the $\delta(z)$ probability for a SN to be
young, since the two effects play their roles, but} the drift is negligible
within the SNfactory dataset ($\Delta\delta(z)<3\%$ from the highest to the
lowest redshift of the dataset)\nn{; thus} one can readily use $y^i = p_y^i$ to
minimize Eq.~\ref{eq:likelihood} with respect to $\vec{\theta}$. Results on
fitting the SNfactory SNe with this model are shown Table~\ref{tab:modelresults}
and illustrated in Fig.~\ref{fig:stretchlssfr}. \yc{\emph{Now I'm confused why
one does not have \textbf{exactly} $y^i = p_y^i$ when LsSFR measurement is
available, but you seem to say it's an assumption.}}

@@

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch (expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances.
We see in this figure that the measured mean SN~Ia stretch per redshift bins of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-hist_on.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift. Markers show the mean stretch measured in
        redshift bins of equal sample size, indicated in light gray at the
        bottom of each redshift bins. Full and light markers are used when
        considering the fiducial or the conservative samples, respectively. The
        orange horizontal line represents the mean stretch of the fiducial
        sample, illustrating the expectation if the SN stretch distribution is
        not evolving with redshift. Best fits of our Base drifting model are
        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
        the conservative one or the SNfactory dataset only, respectively; all
        are compatible. The light-blue band illustrates the amplitude of the
        error (incl. covariance) of the best fit model when considering the
        fiducial dataset. The redshift distributions of each sample are
        shown as light-histograms, see lower left legend. \yc{\emph{This is upper panel of Fig.3? Why is it again split in subsamples, while it was not deemed necessary in dedicated Fig.3?}}}
    \label{fig:modelall}
\end{figure*}

@@

Since the various models have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes extra degrees of
freedom to avoid over-fitting the data, and is defined as follow:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k
\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model is the one
with the smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
a Gaussian probability distribution. However, we highlight that the best of the
compared models \nn{is not meant to} be the true underlying model. \nn{It only
means that it is the best at describing the observations while minimizing the
number of free parameters}. \yc{\emph{why this sentence? Is there any
\textbf{true} intrinsic model?}}

@@

Remark that, when comparing models adjusted on individual subsamples rather than
globally, the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
k\ln(n)$, with $n$ the number of data points) might be better suited than AIC,
since it explicitly accounts for the fact that each subsample is fitted
separately: the sample-based model BIC is rightfully the sum of the BIC for each
sample. Doing so, we find $\Delta\mathrm{BIC}=-48$, again refuting the
sample-based asymmetric Gaussian model as being as pertinent as the Base model.

@@

The amplitude of this magnitude bias for cosmology is beyond the scope of this
paper given the complexity of the BBC analysis. It would require a full study
using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
asymmetric modeling as part of the BBC simulations. However, we already
highlight that even if a non-drifting sample-based model could provide
comparable result in the volume-limited part of the various samples, these
models would differ when extrapolating at higher redshifts, precisely where the
underlying distribution will matter for correcting Malmquist biases.

In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
stress that correctly modeling of potential SN redshift drift should be further
studied and care should be made when using samples affected by selection
effects.

@@

In addition to this ``base'' modeling, we have studied various distributions,
including redshift independent models; we also studied the prediction from the
``Beams with Bias Correction'', a Malmquist bias correction algorithm used by
several recent SN~Ia cosmological analyses, implemented as a per-sample
asymmetric Gaussian \nn{stretch} distribution.

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is significantly redshift
        dependent, as previously suggested by e.g.~\cite{howell2007}. This
        result is largely independent of details on each age-population
        modeling.

    \item Redshift-independent models are indeed excluded as suitable
        descriptions of the data relative to our Base model. This model assumes
        that: (1) the younger population has a unimodal Gaussian stretch
        distribution, while the older population stretch distribution is
        bimodal, one mode being the same as the young one; (2) the evolution of
        the relative fraction of younger and older SNe~Ia follows the prediction
        made in \cite{rigault2018}. This second result strongly supports the
        existence of both young and old SN~Ia populations, in agreement with
        rate studies. \yc{\emph{ref???}}

    \item \nn{Models} using survey-based asymmetric Gaussian distributions, as
        done, e.g., in the current implementation of the BBC, are excluded to be
        good descriptions of the data relative to our drifting model. Hence,
        the sample-based approach does not accurately account for redshift drift
        and even less so as survey span increasingly larger redshift ranges. We
        stipulate that, even if extra degrees of freedom might be acceptable
        given the large number of SNe~Ia in cosmological studies, extrapolating
        the SN property distributions from the volume-limited part of a survey
        to its Malmquist-biased magnitude-limited one  would still be inaccurate
        because of the redshift evolution.

    \item Given the current dataset, we suggest the use of the following stretch
        population model as a function of redshift:
        \begin{align*}
        \label{eqconclusion:stretchz}
            X_1\left(z \right) =
            \delta(z)&\times\mathcal{N}(\mu_1,\sigma_1{}^2)\,+\nonumber\\
            (1-\delta(z))&\times \left[a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
            (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2)\right]
            \tag{\ref{eq:stretchz}}
        \end{align*}
        with $a=0.51$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
        $\sigma_2=0.56$ (see Table~\ref{tab:modelresults}), and using the
        age-population drift model         \begin{align*}
            \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
            \tag{\ref{eq:delta}}
        \end{align*}
        with $K=0.87$.
\end{enumerate}

In this paper, we considered a simple Gaussian mixture modeling, but additional
data free from significant Malmquist bias would enable us to refine it as
necessary. We note that samples at the low- and high-redshift ends of the Hubble
diagram would be particularly helpful for this drifting analysis; fortunately
this will soon be provided by the Zwicky Transient Facility \citep{bellm2019,
graham2019} \nn{on one hand}, and Subaru and SeeChange SNe~Ia programs \nn{on
the other hand}, respectively. \nnn{need to check which is which}
